Evidence for a Color Parameter Unassociated With Dust Within the Type~Ia Supernovae of the Nearby Supernova Factory

### 2016-11-08 apj.tex
finishing up Gerard's comments

In this article we use the idea that spectral indicators carry information on intrinsic supernova colors at peak magnitude.
We allow for three independent color parameters, one attributed to intrinsic supernova
properties and two  attributed to
extrinsic processes, which we attribute to dust.  The data used in this analysis are described in \S\ref{data:sec}.
In \S\ref{model:sec} we present a
first analysis with two extrinsic parameters and omitting the intrinsic parameter.  The properties of the two extrinsic parameters
are consistent with the expectations of parameterized dust designed for the Milky Way, and do an excellent job of fitting the data of the out-of-sample
supernova SN2014J that has an extreme total-to-selective extinction.
In \S\ref{model2:sec} we add to analysis the intrinsic parameter, and find strong evidence for its influence on observed magnitudes.
We summarize our findings, and relate our results with light-curve shape, SiII velocity, and the ``mass step'' in SN~Ia Hubble
residuals in \S\ref{discussion:sec}.

@@

@@

\color{red}
Recall that our model does not specify the color of a supernova that suffers no extrinsic effects, but rather sets the
means of $k_0$ and $k_1$ to zero.
There are unconstrained $k_0^0$, $k_1^0$ that correspond to an intrinsic supernova.  The true extinction is
$A^T_X(k_0,k_1)= A_X(k_0,k_1) -  A_X(k^0_0,k^0_1)$ where $A_X(k_0,k_1) =  \gamma^0_X k_0 + \gamma^1_X k_1$.
The true color excess is $E^T(B-V) = A^T_B-A^T_V$.
Despite this ambiguity, a linear slope relating the $E^T(B-V)$ and $A^T_V$ for a set of supernova is the same slope
relating $A_V$ and $E(B-V)$.
\color{black}
Figure~\ref{avebv:fig} shows the relationship between $A_V$ and $E(B-V)$, from
which a linear fit gives a slope of $R_V=2.77^{+0.08}_{-0.09}$.
\color{red}
The full set of supernovae does not coincide with the expectations of a single linear law; the supernova population rather appears to be peaked
on a ridge line, with a tail distributed below.  That is, for a fixed color excess, the amount of $V$-band extinction is bounded above
but can have values that extend toward lower values.

@@

\color{red}
To help visualize the above result, Figure~\ref{plane:fig} shows two perspectives in $UVI$-space
of the vectors corresponding to $\gamma^i_X/(\gamma^i_B-\gamma^i_V)$ of our model,
and $a(X)$, $b(X)$ of the \citetalias{1999PASP..111...63F} model.
The left plot shows that none of the four vectors are closely aligned with another, while the right plot shows that
the planes defined by our model (solid lines) and the  \citetalias{1999PASP..111...63F} model (dashed lines) are nearly coplanar.
If $R^F_V=2.4$ were constant in our sample, then $A_X = A_V\left(a(X) + \frac{b(X)}{2.4}\right)$;  this combination of $a$ and $b$ is
almost perfectly aligned with $\gamma^0_X/(\gamma^0_B-\gamma^0_V)$, as shown in the dot-dashed line.  
The $\gamma^1$ terms capture the remaining magnitude variations due to the fact that $R^F_V$ is not constant in the sample.

@@

@@

>>>>

In this article we use the idea that spectral indicators carry information on intrinsic supernova colors at peak magnitude.
\color{red}
This approach is taken by \citet{2011A&A...529L...4C}, who find that after standardization based on Ca and Si features, remaining residual color
variation is consistent with Milky Way dust laws.
We accommodate  three independent color parameters: the model is constructed such that one is attributed to intrinsic supernova
properties and the two others to
extrinsic processes that are attributed to dust.
\color{black}
The data used in this analysis are described in \S\ref{data:sec}.
In \S\ref{model:sec} we present a
first analysis with two extrinsic parameters and omitting the intrinsic parameter.  The properties of the two extrinsic parameters
are consistent with the expectations of parameterized dust designed for the Milky Way, and do an excellent job of fitting the data of the out-of-sample
supernova SN2014J that has an extreme total-to-selective extinction.
In \S\ref{model2:sec} we add to analysis the intrinsic parameter, and find strong evidence for its influence on observed magnitudes.
We summarize our findings, and relate our results with light-curve shape, SiII velocity, and the ``mass step'' in SN~Ia Hubble
residuals in \S\ref{discussion:sec}.

@@

\color{red}
For $N$ supernovae there are $8N$ observables.  For the top-level model parameters, there are $3N$ spectral parameters, $2(N-2)$
color parameters, $(N-1)$ magnitude offsets,  $5 \times 6$ global coefficients, and $10$ parameters that describe the intrinsic
$C_c$.  For $N=168$ supernovae, there are 1344 observables and 1043 top-level parameters.
\color{black}

@@

\color{red}
Recall that our model does not specify the color of a supernova that suffers no extrinsic effects, but rather sets the
means of $k_0$ and $k_1$ to zero.
There are unconstrained $k_0^0$, $k_1^0$ that correspond to an intrinsic supernova.  The true extinction is
$A^T_X(k_0,k_1)= A_X(k_0,k_1) -  A_X(k^0_0,k^0_1)$ where $A_X(k_0,k_1) =  \gamma^0_X k_0 + \gamma^1_X k_1$.
The true color excess is $E^T(B-V) = A^T_B-A^T_V$.
Despite this ambiguity, a linear slope relating the $E^T(B-V)$ and $A^T_V$ for a set of supernova is the same slope
relating $A_V$ and $E(B-V)$.
\color{black}
Figure~\ref{avebv:fig} shows the relationship between $A_V$ and $E(B-V)$, from
which a linear fit gives a slope of $R_V=2.77^{+0.08}_{-0.09}$.
\color{red}
The slope uncertainties are the $68$\% credible interval from the Bayesian analysis and does not incorporated goodness
 of fit.  Indeed,
the full set of supernovae does not coincide with the expectations of a single linear law; the supernova population rather appears to be peaked
on a ridge line, with a tail distributed below.  That is, for a fixed color excess, the amount of $V$-band extinction is bounded above
but can have values that extend toward lower values.

@@

\color{red}
To help visualize the above result, Figure~\ref{plane:fig} shows two perspectives in $UVI$-space
of the vectors corresponding to $\gamma^i_X/(\gamma^i_B-\gamma^i_V)$ of our model,
and $a(X)$, $b(X)$ of the \citetalias{1999PASP..111...63F} model.
The left plot shows that none of the four vectors are closely aligned with another, while the right plot shows that
the planes defined by our model (solid lines) and the  \citetalias{1999PASP..111...63F} model (dashed lines) are nearly coplanar.
If $R^F_V=2.4$ were constant in our sample, then $A_X = A_V\left(a(X) + \frac{b(X)}{2.4}\right)$;  this combination of $a$ and $b$ is
almost perfectly aligned with $\gamma^0_X/(\gamma^0_B-\gamma^0_V)$, as shown in the dot-dashed line.  
The $\gamma^1$ terms capture the remaining magnitude variations due to the fact that $R^F_V$ is not constant in the sample:
if the $\gamma$'s are due to  \citetalias{1999PASP..111...63F} dust, the only way our analysis could recover
both $\gamma^0$ and $\gamma^1$ is for there to be variation in $R^F_V$.

@@

\color{red}
For $N$ supernovae there are $8N$ observables.  For the top-level model parameters, there are $3N$ spectral parameters, $3(N-2)$
color parameters, $(N-1)$ magnitude offsets,  $5 \times 7$ global coefficients, and $10$ parameters that describe the intrinsic
$C_c$.  For $N=168$ supernovae, there are 1344 observables and 1214 top-level parameters.
\color{black}

@@

\color{red}
The per-supernova color parameters are not strongly correlated,
\begin{equation}
Cor(k_0, k_1, D) =
\begin{pmatrix}
\begin{array}{rrr}
1.00 & 0.06 & -0.01 \\
0.06 & 1.00 & 0.00 \\
-0.01 & 0.00 & 1.00
\end{array}
\end{pmatrix}.
\end{equation}
In particular, the Pearson correlation coefficients between the ``intrinsic'' $D$ and ``extrinsic'' $k$ parameters are $-0.01$ and $0.00$.
So while in principle a single physical color effect can affect both intrinsic and extrinsic parameters, from the correlations
there is no evidence
that this occurs in our fits.

The distributions for $\Delta$, $EW_{Ca}$, $EW_{Si}$, $\lambda_{Si}$, $E_{\gamma^0}(B-V)$, 
$E_{\gamma^1}(B-V)$, and
$E_{\delta}(B-V)$ for all Monte Carlo links for all supernovae are shown in Figure~\ref{perobject2:fig}.

\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=5.5in]{output15/perobject_corner.pdf} 
   \caption{
   \color{red}
   Distributions for the supernova parameters $\Delta$, $EW_{Ca}$, $EW_{Si}$, $\lambda_{Si}$, $E_{\gamma^0}(B-V)$,  $E_{\gamma^1}(B-V)$,  and $E_{\delta}(B-V)$, as well as the grey offset
$\Delta$.  All Monte Carlo links are plotted, so that each supernova contributes a cloud of points.
\color{black}
   \label{perobject2:fig}}
\end{figure}

Our code is validated
though the analysis of a simulated dataset: the input parameters are recovered within  uncertainties.
\color{black}
